---
title: "EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r library}
library(corrplot)
library(ggplot2)
library(magrittr)
library(tibble)
library(ggcorrplot)
library(tidyverse)
library(ISLR)
```


Load in the data.

```{r}
#db_binary <- read.csv("~/Desktop/STA 325/diabetesprediction/data/diabetes_binary.csv")
#db_5050 <- read.csv("~/Desktop/STA 325/diabetesprediction/data/diabetes_binary_5050split.csv")
#db_012 <- read.csv("~/Desktop/STA 325/diabetesprediction/data/diabetes_012.csv")

db_binary <- read_csv("diabetes_binary.csv")
db_5050 <- read_csv("diabetes_binary_5050split.csv")
db_012 <- read_csv("diabetes_012.csv")

# Making all categorical variables into factors
# cols <- names(db_012) %in% c("BMI", "MentHlth", "PhysHlth", "Age")
# db_012[!cols] <- data.frame(lapply(db_012[!cols], factor))
# tibble(db_012)
```

We can begin by creating correlation matrix plots to assess if there are any predictors with significant correlation towards our response variable `Diabetes_012`. There are 21 predictors, and therefore, creating one correlation matrix plot would be busy and difficult to understand. Since we are using a correlation matrix plot to investigate the predictors' correlation with the response - and not one another - splitting the variables into three visualizations is appropriate.

```{r corplots}
first_vars <- db_012[1:8]
second_vars <- db_012[c(1,9:15)]
third_vars <- db_012[c(1,16:22)]

# model.matrix(Diabetes_012 ~ ., data=first_vars) %>% 
#   cor(use="pairwise.complete.obs") %>% 
#   ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=2)
# 
# model.matrix(Diabetes_012 ~ ., data=second_vars) %>% 
#   cor(use="pairwise.complete.obs") %>% 
#   ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=2)
# 
# model.matrix(Diabetes_012 ~ ., data=third_vars) %>% 
#   cor(use="pairwise.complete.obs") %>% 
#   ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=2)

corrplot.mixed(cor(first_vars),
               lower = "number", 
               upper = "circle", 
               tl.pos = "lt")
corrplot.mixed(cor(second_vars),
               lower = "number", 
               upper = "circle", 
               tl.pos = "lt")
corrplot.mixed(cor(third_vars),
               lower = "number", 
               upper = "circle", 
               tl.pos = "lt")

# corrplot(cor(db_012[1:8]))
# corrplot(cor(db_012[c(1,9:15)]))
# corrplot(cor(db_012[c(1,16:22)]))
```

From the correlation matrix plots, it seems that there are 12 significant main effects on our response variable `Diabetes_012`. There are nine variables (`HighBP`, `HighChol`, `BMI`, `Stroke`, `HeartDiseaseorAttack`,  `GenHlth`, `PhysHlth`, `DiffWalk`, `Age`) that demonstrate a significant positive correlation with our response, while there are three variables (`PhysActivity`, `Education`, `Income`) with a significant negative correlation.

We should validate these assumptions from the correlation matrix plots by developing additional exploratory visualizations between our predictors of interest and the response. 

```{r viz}
bmi_plot <- ggplot(db_012, aes(factor(Diabetes_012), BMI)) +
  geom_boxplot() +
  labs(title = 'Diabetes And BMI', x = "Diabetes_012", y = "BMI")
bmi_plot

bp_plot <- ggplot(db_012, aes(x = factor(Diabetes_012), fill = factor(HighBP))) +
  geom_bar(position = "fill") +
  labs(title = 'Diabetes And HighBP', x = "Diabetes_012", y = "Count", 
       fill = "HighBP") +
    scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF"))
bp_plot

chol_plot <- ggplot(db_012, aes(x = factor(Diabetes_012), fill = factor(HighChol))) +
  geom_bar(position = "fill") +
  labs(title = 'Diabetes And HighChol', x = "Diabetes_012", y = "Count", 
       fill = "HighChol") +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF"))
chol_plot

stroke_plot <- ggplot(db_012, aes(x = factor(Diabetes_012), fill = factor(Stroke))) +
  geom_bar(position = "fill") +
  labs(title = 'Diabetes And Stroke', x = "Diabetes_012", y = "Count", 
       fill = "Stroke") +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF"))
stroke_plot

heart_plot <- ggplot(db_012, aes(x = factor(Diabetes_012), fill = factor(HeartDiseaseorAttack))) +
  geom_bar(position = "fill") +
  labs(title = 'Diabetes And Heart Disease or Attack', x = "Diabetes_012", y = "Count", 
       fill = "Heart Disease or Attack") +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF"))
heart_plot

genhealth_plot <- ggplot(db_012, aes(factor(Diabetes_012), GenHlth)) +
  geom_boxplot() +
  labs(title = 'Diabetes And General Health', x = "Diabetes_012", y = "General Health")
genhealth_plot

physhealth_plot <- ggplot(db_012, aes(factor(Diabetes_012), PhysHlth)) +
  geom_boxplot() +
  labs(title = 'Diabetes And Physical Health', x = "Diabetes_012", y = "Physical Health")
physhealth_plot

diffwalk_plot <- ggplot(db_012, aes(x = factor(Diabetes_012), fill = factor(DiffWalk))) +
  geom_bar(position = "fill") +
  labs(title = 'Diabetes And Difficulty Walking', x = "Diabetes_012", y = "Count", 
       fill = "Difficulty Walking") +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF"))
diffwalk_plot

age_plot <- ggplot(db_012, aes(factor(Diabetes_012), Age)) +
  geom_boxplot() +
  labs(title = 'Diabetes And Age', x = "Diabetes_012", y = "Age")
age_plot
```

