---
title: "EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r library}
library(corrplot)
library(ggplot2)
library(magrittr)
library(tibble)
library(ggcorrplot)
library(tidyverse)
library(ISLR)
require(nnet)
library(car)
library(glmnet)
library(pROC)
library(pscl)
```


Load in the data.

```{r data}
db_binary <- read_csv("diabetes_binary.csv")
db_5050 <- read_csv("diabetes_binary_5050split.csv")
db_012 <- read_csv("diabetes_012.csv")
```

```{r wrangling}
# Making all categorical variables into factors
cols <- names(db_012) %in% c("Age", "BMI", "MentHlth", "PhysHlth", "Education")
db_binary[!cols] <- data.frame(lapply(db_binary[!cols], factor))
# Coercing `PhysHlth` and "MentHlth` to integer type
db_binary["PhysHlth"] <- data.frame(lapply(db_binary["PhysHlth"], as.integer))
db_binary["MentHlth"] <- data.frame(lapply(db_binary["MentHlth"], as.integer))

# Checking to see if `Age` groups can be consolidated
age_check <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, Age), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Age vs. Diabetes Diagnosis")
age_check 

# Binning `Age` into less more understandable categories
db_binary <- db_binary %>%
  mutate(Age = case_when(
    Age <= 4 ~ "18-39" ,
    Age > 4 & Age <=8 ~ "40-59" ,
    Age > 8 ~ "60+"
  ))

# Converting `Age` to a Factor
db_binary["Age"] <- data.frame(lapply(db_binary["Age"], as.factor))

# Checking to see if `Education` groups can be consolidated
edu_check <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, Education), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Education vs. Diabetes Diagnosis")
edu_mosaic

# Binning `Education` into less more understandable categories
db_binary <- db_binary %>%
  mutate(Education = case_when(
    Education <= 3 ~ "Some HighSchool or Less" ,
    Education == 4 ~ "High School Grad" ,
    Education == 5 ~ "Some College",
    Education == 6 ~ "College Graduate"
  ))

# Converting `Education` to a Factor
db_binary["Education"] <- data.frame(lapply(db_binary["Education"], as.factor))

db_binary <- db_binary %>%
  mutate(Education = fct_relevel(Education,"Some HighSchool or Less" ,"High School Grad", "Some College","College Graduate"))

```

We can begin by creating correlation matrix plots to assess if there are any predictors with significant correlation towards our response variable `Diabetes_012`. There are 21 predictors, and therefore, creating one correlation matrix plot would be busy and difficult to understand. Since we are using a correlation matrix plot to investigate the predictors' correlation with the response - and not one another - splitting the variables into three visualizations is appropriate.

```{r mosaic plots}
# HighBP
bp_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, HighBP), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "High BP vs. Diabetes Diagnosis")
bp_mosaic

# HighChol
chol_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, HighChol), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "High Cholesterol vs. Diabetes Diagnosis")
chol_mosaic

# CholCheck ---> not valuable, skewed
chol_check_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, CholCheck), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Cholesterol Check vs. Diabetes Diagnosis")
chol_check_mosaic

#BMI
bmi_df <- db_binary %>%
  group_by(Diabetes_binary) %>%
  summarize(median=median(BMI))

bmi_mosaic <- ggplot(data = db_binary, mapping = aes(x = BMI, colour = Diabetes_binary)) +
  geom_freqpoly(binwidth = 1, aes(fill = Diabetes_binary)) +
  geom_vline(data = bmi_df, aes(xintercept = median, color = Diabetes_binary),
                                   linetype = "dotted") + 
  labs(title = "Body Mass Index vs. Diabetes Diagnosis") +
  scale_colour_manual(values = c("#7EAED2FF", "#EDB687FF"))
bmi_mosaic 

# Smoker --> not particulary significant
smoker_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, Smoker), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Smoker vs. Diabetes Diagnosis")
smoker_mosaic

# Stroke --> too skewed
stroke_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, Stroke), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Stroke vs. Diabetes Diagnosis")
stroke_mosaic

# HeartDiseaseorAttack --> too skewed?
heart_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, HeartDiseaseorAttack), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Heart Disease/Attack vs. Diabetes Diagnosis")
heart_mosaic

# PhysActivity --> kind of close?
phys_act_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, PhysActivity), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Physical Activity vs. Diabetes Diagnosis")
phys_act_mosaic

# Fruits --> very close
fruit_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, Fruits), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Fruits vs. Diabetes Diagnosis")
fruit_mosaic

# Veggies --> very close + skewed
veg_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, Veggies), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Veggies vs. Diabetes Diagnosis")
veg_mosaic

# HvyAlcoholConsump --> very skewed
alc_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, HvyAlcoholConsump), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Heavy Alcohol Consumption vs. Diabetes Diagnosis")
alc_mosaic

# AnyHealthcare --> very skewed
care_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, AnyHealthcare), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Healthcare vs. Diabetes Diagnosis")
care_mosaic

# NoDocbcCost -->skew
doc_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, NoDocbcCost), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "No Doctor/Cost vs. Diabetes Diagnosis")
doc_mosaic

# GenHlth
gen_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, GenHlth), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "General Health vs. Diabetes Diagnosis")
gen_mosaic

# MentHlth --> same median, not applicable
ment_df <- db_binary %>%
  group_by(Diabetes_binary) %>%
  summarize(median=median(MentHlth))

ment_mosaic <- ggplot(data = db_binary, mapping = aes(x = MentHlth, colour = Diabetes_binary)) +
  geom_freqpoly(binwidth = 1) +
  geom_vline(data = ment_df, aes(xintercept = median, color = Diabetes_binary),
                                  linetype = "dotted") + 
  labs(title = "Mental Health vs. Diabetes Diagnosis", x = "Number of Days") +
  scale_colour_manual(values = c("#7EAED2FF", "#EDB687FF"))
ment_mosaic 

# PhysHlth --> maybe?

phys_df <- db_binary %>%
  group_by(Diabetes_binary) %>%
  summarize(median=median(PhysHlth))

phys_mosaic <- ggplot(data = db_binary, mapping = aes(x = PhysHlth, colour = Diabetes_binary)) +
  geom_freqpoly(binwidth = 1) +
  geom_vline(data = phys_df, aes(xintercept = median, color = Diabetes_binary),
                                  linetype = "dotted") + 
  labs(title = "Physical Health vs. Diabetes Diagnosis", x = "Number of Days") +
  scale_colour_manual(values = c("#7EAED2FF", "#EDB687FF"))
phys_mosaic 

# DiffWalk --> maybe?
walk_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, DiffWalk), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Difficulty Walking vs. Diabetes Diagnosis")
walk_mosaic

# Sex --> little difference
sex_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, Sex), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Sex vs. Diabetes Diagnosis")
sex_mosaic

# Age 
age_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, Age), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Age vs. Diabetes Diagnosis")
age_mosaic

# Education --> maybe?
edu_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, Education), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Education vs. Diabetes Diagnosis") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
edu_mosaic

# Income
inc_mosaic <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Diabetes_binary, Income), fill = Diabetes_binary)) +
  scale_fill_manual(values = c("#7EAED2FF", "#EDB687FF")) +
  labs(title = "Income vs. Diabetes Diagnosis")
inc_mosaic
```


Good: HighBP, HighChol, BMI, GenHlth, Age, Education, Income
Maybe: PhysActivity(slightly skewed),  PhysHlth (skewed but diff), DiffWalk (skewed but diff)
Bad: CholCheck (too skewed, little obs.), Smoker(very even), Stroke (too skewed, little obs.), HeartDiseaseorAttack (skewed, little obs.), Fruits(very even), Veggies(skewed and little difference), HvyAlcoholConsumpt(very skewed, little diff), AnyHealthcare(very skewed little difference), NoDocbcCost(very skewed, little diff), MentHlth(no diff in median, both 0), Sex(little diff)

From the mosaic plots, it seems that there are 10 significant main effects on our response variable `Diabetes_binary`. These variables (`HighBP`, `HighChol`, `BMI`, `GenHlth`, `Age`, `Education`, `Income`, `PhysActivity`, `PhysHlth`, `DiffWalk`) all demonstrated that the proportion of respondents with a diabetes diagnosis considerably changed when accounted for, meaning it is likely they share some relationship with our response.

The other 11 predictors in our data set do not appear to share a meaningful association with `Diabetes_binary`. Several of the predictors (`CholCheck`, `Stroke`, `HeartDiseaseorAttack`) were skewed towards their baseline (ie. does not apply to the respondent), and therefore, they provide limited observations of individuals having said condition, behavior, or circumstance. Additionally, the variables `Smoker`, `HvyAlcoholConsumption`, `Fruits`, `Veggies`, `HeartDiseaseorAttack`, `AnyHealthcare`, `NoDocbcCost`, `MentHlth`, and `Sex` exhibited little difference in the proportion of respondents of having diabetes vs. not having diabetes when depended on. This indicates that they may have a very weak or no correlation with the response `Diabetes_binary`.

Seeing as the primary focus of our analysis is to identify particular demographics and behaviors that are associated with diabetes, and in turn, make tangible recommendations to prevent its incidence, we find it justifiable to remove these variables from the model building process. This will allow us to construct a more interpretable model that promotes our analysis goals. 

We should validate these assumptions from the correlation matrix plots by developing additional exploratory visualizations between our predictors of interest and the response. 

```{r interactions}
bp_phys_mos <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(HighBP,Diabetes_binary, PhysActivity), fill = HighBP, alpha = Diabetes_binary)) + 
  scale_alpha_manual(values =c(.5,.9)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) + 
  labs(y="Diabetes Diagnosis", x="HighBP:PhysActivity", title = "Interaction Effect Between High BP and Physical Activity ")
bp_phys_mos

hlth_age_mos <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(GenHlth,Diabetes_binary, Age), fill = GenHlth, alpha = Diabetes_binary)) + 
  scale_alpha_manual(values =c(.5,.9)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) + 
  labs(y="Diabetes Diagnosis", x="Age:GenHlth", title = "Interaction Effect Between Age and General Health")
hlth_age_mos

bp_age_mos <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(HighBP,Diabetes_binary, Age), fill = HighBP, alpha = Diabetes_binary)) + 
  scale_alpha_manual(values =c(.5,.9)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) + 
  labs(y="Diabetes Diagnosis", x="HighBP:Age", title = "Interaction Effect Between High BP and Age")
bp_age_mos

edu_inc_mos <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Education,Diabetes_binary, Income), fill = Education, alpha = Diabetes_binary)) + 
  scale_alpha_manual(values =c(.5,.9)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) + 
  labs(y="Diabetes Diagnosis", x="Education:Income", title = "Interaction Effect Between Education and Income")
edu_inc_mos

age_diff_mos <- ggplot(data = db_binary) +
  geom_mosaic(aes(x=product(Age,Diabetes_binary, DiffWalk), fill = Age, alpha = Diabetes_binary)) + 
  scale_alpha_manual(values =c(.5,.9)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5)) + 
  labs(y="Diabetes Diagnosis", x="Age:DiffWalk", title = "Interaction Effect Between Age and Difficulty Walking")
age_diff_mos
```

Next, we will evaluate whether our continuous variable, BMI, is skewed:

```{r hist}
ggplot(db_012, aes(x = BMI)) + 
  geom_histogram(fill = "blue", binwidth = 2)
```

```{r dset-glimpse}
glimpse(db_binary)
```

```{r starter-log-model}
log_model_1 <- glm(Diabetes_binary ~ ., data = db_binary, family = "binomial")
summary(log_model_1)
```

```{r interactions-log-model}
log_model_2 <- glm(Diabetes_binary ~ HighBP + HighChol + CholCheck + BMI + Stroke + HeartDiseaseorAttack + PhysActivity + Fruits + HvyAlcoholConsump + GenHlth + MentHlth + DiffWalk + Sex + Age + Education + Income + Age * HighBP + Age * HeartDiseaseorAttack + Age * HighChol + Age * GenHlth + Sex * GenHlth, data = db_binary, family = "binomial")
summary(log_model_2)
```

```{r larger-interactions-log-model}
log_model_3 <- glm(Diabetes_binary ~ HighBP + HighChol + CholCheck + BMI + Stroke + HeartDiseaseorAttack + PhysActivity + Fruits + HvyAlcoholConsump + GenHlth + MentHlth + DiffWalk + Sex + Age + Education + Income + Age * HighBP + Age * HeartDiseaseorAttack + Age * HighChol + Age * GenHlth + Sex * GenHlth + Sex * HighBP + Sex * MentHlth, data = db_binary, family = "binomial")
summary(log_model_3)
```

```{r log-model-4}
log_model_4 <- glm(Diabetes_binary ~ HighBP + HighChol + BMI + GenHlth + PhysHlth + DiffWalk + Age + Education + Income + Age * HighBP + Age * HighChol + GenHlth * PhysHlth + Age * DiffWalk, data = db_binary, family = "binomial")
summary(log_model_4)
```

```{r multi-collinearity}
vif(log_model_1)
vif(log_model_2)
vif(log_model_3)
```

```{r anova}
anova(log_model_2, log_model_3, test ="Chisq")
#anova(log_model_1)
#anova(log_model_2)
#anova(log_model_3)
```
```{r tr-te-split}
split1 <- sample(c(rep(0, 0.7 * nrow(db_binary)), rep(1, 0.3 * nrow(db_binary))))
train <- db_binary[split1 == 0, ]
test <- db_binary[split1== 1, ]
```

```{r lasso}
x = data.matrix(select(train, -Diabetes_binary))
y = data.matrix(train$Diabetes_binary)
#cv.lasso <- cv.glmnet(x, y, alpha = 1, family = "binomial")
#plot(cv.lasso)
```

```{r lasso-model}
model <- cv.glmnet(x, y, alpha = 1, family = "binomial")
#model <- glmnet(x, y, alpha = 1, family = "binomial", lambda = cv.lasso$lambda.min)
coef(model)
```

```{r roc}
x2 = data.matrix(select(test, -Diabetes_binary))
prob = predict(model, x2, type=c("response"))
g <- roc(test$Diabetes_binary ~ prob, plot = TRUE, print.auc = TRUE)
plot(g)
```

```{r misclassification}
glm.predict = ifelse(prob>0.5, 1, 0)
mean(glm.predict != test$Diabetes_binary)
table(glm.predict, test$Diabetes_binary)
```

```{r gam}
log_model_orig <- glm(Diabetes_binary ~ HighBP + HighChol + CholCheck + Stroke + HeartDiseaseorAttack + PhysActivity + Fruits + HvyAlcoholConsump + GenHlth + MentHlth + DiffWalk + Sex + Age + Education + Income + Age * HighChol + Age * GenHlth + Sex * GenHlth + Sex * HighBP, data = db_binary, family = "binomial")
summary(log_model_comp)

log_model_comp <- glm(Diabetes_binary ~ HighBP + HighChol + CholCheck + BMI + Stroke + HeartDiseaseorAttack + PhysActivity + Fruits + HvyAlcoholConsump + GenHlth + MentHlth + DiffWalk + Sex + Age + Education + Income + Age * HighChol + Age * GenHlth + Sex * GenHlth + Sex * HighBP, data = db_binary, family = "binomial")
summary(log_model_comp)

gam.1 <- gam(Diabetes_binary ~ HighBP + HighChol + CholCheck + s(BMI, df=5) + Stroke + HeartDiseaseorAttack + PhysActivity + Fruits + HvyAlcoholConsump + GenHlth + MentHlth + DiffWalk + Sex + Age + Education + Income + Age * HighChol + Age * GenHlth + Sex * GenHlth + Sex * HighBP, data = db_binary, family = "binomial")
summary(gam.1)
plot(gam.1, se = TRUE, col = "blue")
anova(log_model_orig, log_model_comp, gam.1)
```

